Kepler 1627 Ab: A 38$\,$Myr Mini-Neptune in the Kepler Field

### 2021-08-24 paper/ms.tex
update ms

@@

\title{
  An Adolescent Mini-Neptune in the Kepler Field
}

@@

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Combining
  kinematic information from Gaia, stellar rotation periods from TESS,
  and Li abundances from ground-based spectra, we show that
  Kepler\,1627 is a member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr
  cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
  with a precise age yet found by the main Kepler mission.  From both
  unresolved starspot-crossing events in the Kepler photometry and
  from a Keck/HIRES transit we find that the planet is
  likely on an orbit nearly aligned with the stellar spin axis.  We
  also find from the short-cadence Kepler photometry that although the
  flare arrival time distribution is probably Poissonian, there may be
  an excess of flares separated by the orbital and synodic periods,
  and linear combinations thereof.  The $\delta$~Lyr cluster is one of
  many stellar groups whose properties have been clarified using the
  Gaia data; many of the other new clusters also overlap with
  confirmed and candidate exoplanets.  Further studies that combine
  Gaia kinematics and TESS rotation periods are poised to expand the
  census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

@@

A previous clustering analysis of the Gaia data reported that
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
\cn\ \citep{kounkel_untangling_2019}.  Given the previous statistical
validation the close-in mini-Neptune Kepler 1627b
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we examined the age of the star and its putative host cluster more
closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
of the cluster.  Focusing on the planet itself,
we confirm that despite the presence of a previously unreported binary companion,
the planet orbits the G-dwarf primary star on an orbit that appears aligned
with the stellar rotation axis
(Section~\ref{sec:planet}).  We conclude by
highlighting broader implications for our ability to age-date a
statistical sample of planets (Section~\ref{sec:conc}).

@@

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35^{+6}_{-5}$\,Myr old.}  {\it Top:}
    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
    the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
    field (gray).  The right panel provides a zoomed
    view of stars on the pre-main-sequence.  The \cn\ and IC\,2602 are
    approximately the same isochronal age.  {\it Bottom:} TESS and
    Kepler stellar rotation periods versus dereddened Gaia colors, with
    the Pleiades and Praesepe (650\,Myr) shown for
    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.
    The \cn\ is gyrochronally younger than the Pleiades.
   \label{fig:age}
	}
\end{figure*}

@@

The ages returned by this procedure depend on the ages assumed for
each reference cluster.  We adopted a 115\,Myr age for the Pleiades,
and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.
The age of IC\,2602 however is the most important
ingredient, since it receives the most weight in the interpolation.
Plausible ages for IC\,2602
span 30\,Myr to 46\,Myr, with slightly older ages being preferred by
the lithium-depletion-boundary (LDB) measurements and younger ages by
the main-sequence turn-off
\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019,bouma_cluster_2020}.
If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
be $31^{+5}_{-4}$\,Myr old.  For the converse case of 46\,Myr, the
\cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
35\,Myr age as our reference, yielding an age for the \cn\ of
$35^{+6}_{-5}$\, Myr, though it should be understood based on the
systematic uncertainties that an age of up to $\approx50$\,Myr remains
plausible.  Detailed analyses of the LDB or main-sequence turn-off in
the \cn\ could help resolve the issue, but both are beyond our scope.

@@

In brief, the age of the \cn\ cluster based on the empirical
fitting procedure described in Section~\ref{sec:cmd} is
$35^{+6}_{-5}$\,Myr.  This procedure hinges on the age adopted for
IC\,2602, since it is the closest reference cluster in the
color-magnitude diagram.  Gyrochronologically, while most of the G and
K dwarfs in the \cn\ have not yet converged to a slow sequence, their
rotation distribution and that of the F dwarfs is consistent with the
same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
\citep{douglas_stephanie_t_2021_5131306}.  The most precise age for \sn\ and its
planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's
spatial, kinematic, and rotational connection to this larger ensemble,
and the assumption that the planet formed very quickly after the star.

@@

We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
models for the Rossiter-McLaughlin (RM) effect, and allowed three parameters
to vary: the sky-projected obliquity, the stellar
equatorial velocity, and the Gaussian dispersion of spectral lines.
We also allowed for an optional linear and quadratic trend, and
a white-noise jitter term.  The quadratic model with the RM
effect is shown in Figure~\ref{fig:obliquity}.  We found $\chi^2_{\rm
red} = 5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} =
4.4$ for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
BIC}_{\rm line} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
{\rm AIC}_{\rm line} - {\rm AIC}_{\rm quad} = 20.6$.  While this
provides strong statistical evidence to prefer the quadratic model
over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
is particularly good.  Considering a quadratic trend alone without any
RM effect yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} =
{\rm BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ (similarly $\Delta
{\rm AIC}=5.2$); from a model-comparison standpoint, there is therefore only
weak statistical evidence to prefer the best-fitting ``RM and
quadratic trend'' model over the best-fitting ``quadratic trend only''
model.  Given the expected semi-amplitude of the velocity anomaly
($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual achieved
precision, the claim of having detected the Doppler
transit is therefore marginal.  
From the MUSCAT3 data however, we do know that the transit was observed, and the
increase in radial velocity seen shortly after ingress is more
consistent with a prograde or polar orbit than a retrograde one.
Fixing the orbit to be either exactly prograde or exactly retrograde
and allowing for a quadratic trend, we find $\Delta {\rm BIC} = {\rm
BIC}_{\rm retrograde} - {\rm BIC}_{\rm prograde} = 20.4$; the
retrograde case is ruled out on this basis.  More precise data will
however still be useful to make a more precise obliquity measurement.

>>>>

\newcommand{\clusterage}{$38^{+6}_{-5}$\,Myr} %

@@

\title{
  %Kepler 1627Ab: A 38 Million Year Old
  Kepler 1627 Ab: A 38$\,$Myr Mini-Neptune in the Kepler Field
}

@@

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Combining
  kinematic information from Gaia, stellar rotation periods from TESS,
  and Li abundances from ground-based spectra, we show that
  Kepler\,1627 is a member of the \clusterage\ old $\delta$~Lyr
  cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
  with a precise age yet found by the main Kepler mission.  From both
  unresolved starspot-crossing events in the Kepler photometry and
  from a Keck/HIRES transit we find that the planet is
  likely on an orbit nearly aligned with the stellar spin axis.  We
  also find from the short-cadence Kepler photometry that although the
  flare arrival time distribution is probably Poissonian, there may be
  an excess of flares separated by the orbital and synodic periods,
  and linear combinations thereof.  The $\delta$~Lyr cluster is one of
  many stellar groups whose properties have been clarified using the
  Gaia data; many other new clusters also overlap with confirmed and
  candidate exoplanets.  Further studies that combine Gaia kinematics
  and TESS rotation periods are poised to expand the census of
  age-dated planets, and our understanding of planetary evolution.
\end{abstract}

@@

A previous clustering analysis of the Gaia data reported that
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
\cn\ \citep{kounkel_untangling_2019}.  Given the previous statistical
validation the close-in mini-Neptune Kepler 1627b
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we examined the age of the star and its putative host cluster more
closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
cluster is \clusterage\ old, and that \sn\ is indeed a member
of the cluster.  Focusing on the planet itself,
we confirm that despite the presence of a previously unreported binary companion,
the planet orbits the G-dwarf primary star on an orbit that appears aligned
with the stellar rotation axis
(Section~\ref{sec:planet}).  We conclude by
highlighting broader implications for our ability to age-date a
statistical sample of planets (Section~\ref{sec:conc}).

@@

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
    the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
    field (gray).  The right panel provides a zoomed
    view of stars on the pre-main-sequence.  The \cn\ and IC\,2602 are
    approximately the same isochronal age.  {\it Bottom:} TESS and
    Kepler stellar rotation periods versus dereddened Gaia colors, with
    the Pleiades and Praesepe (650\,Myr) shown for
    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.
    The \cn\ is gyrochronally younger than the Pleiades.
   \label{fig:age}
	}
\end{figure*}

@@

% Randich+18: 43.7 +4.3 - 3.9 LDB, for E(B-V) = 0.068 +/- 0.025
% Dobbie+10: 46 +6 -5 LDB, for

The ages returned by this procedure depend on the ages assumed for
each reference cluster.  We adopted a 115\,Myr age for the Pleiades,
and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.  The age
of IC\,2602 however is the most important ingredient, since it
receives the most weight in the interpolation.  Plausible ages for
IC\,2602 span 30\,Myr to 46\,Myr, with slightly older ages being
preferred by the lithium-depletion-boundary (LDB) measurements and
younger ages by the main-sequence turn-off
\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019,bouma_cluster_2020}.
If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
be $31^{+5}_{-4}$\,Myr old.  For the converse case of 46\,Myr, the
\cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
38\,Myr age as our reference, yielding an age for the \cn\ of
\clusterage.\footnote{Our own exploration of the PARSEC and MIST
isochrone models over a grid of ages, metallicities, and
reddenings, we yielded the best agreement for this $\approx 38\,$Myr age
as well, given ${\rm [Fe/H] = +0.1}$ and $A_V=0.2$
\citep{bressan_parsec_2012,choi_mesa_2016}.
} While detailed analyses of the LDB or main-sequence turn-off in the
\cn\ could help resolve the tension between the LDB and isochrone
ages, these are beyond our scope.

@@

In brief, the age of the \cn\ cluster based on the empirical
fitting procedure described in Section~\ref{sec:cmd} is
\clusterage.  This procedure hinges on the age adopted for
IC\,2602, since it is the closest reference cluster in the
color-magnitude diagram.  Gyrochronologically, while most of the G and
K dwarfs in the \cn\ have not yet converged to a slow sequence, their
rotation distribution and that of the F dwarfs is consistent with the
same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
\citep{douglas_stephanie_t_2021_5131306}.  The most precise age for \sn\ and its
planet are therefore also \clusterage\ Myr, based on the star's
spatial, kinematic, and rotational connection to this larger ensemble,
and the assumption that the planet formed very quickly after the star.

@@

We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
models for the Rossiter-McLaughlin (RM) effect, and allowed three parameters
to vary: the sky-projected obliquity, the stellar
equatorial velocity, and the Gaussian dispersion of spectral lines.
We also allowed for an optional linear and quadratic trend, and
a white-noise jitter term.  The quadratic model with the RM
effect is shown in Figure~\ref{fig:obliquity}, where the 
prior on $v\sin i$ has been assumed from Table~\ref{tab:starparams}.  We found $\chi^2_{\rm
red} = 5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} =
4.4$ for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
BIC}_{\rm line} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
{\rm AIC}_{\rm line} - {\rm AIC}_{\rm quad} = 20.6$.  While this
provides strong statistical evidence to prefer the quadratic model
over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
is particularly good.  Considering a quadratic trend alone {\it without any
RM effect} yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} =
{\rm BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ (similarly $\Delta
{\rm AIC}=5.2$).  From a model-comparison standpoint, there is therefore only
weak statistical evidence to prefer the best-fitting ``RM and
quadratic trend'' model over the best-fitting ``quadratic trend only''
model.  Given the expected semi-amplitude of the velocity anomaly
($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual achieved
precision, the claim of having detected the Doppler
transit is marginal at best.  
From the MUSCAT3 data however, we do know that the transit was observed, and the
increase in radial velocity seen shortly after ingress is more
consistent with a prograde or polar orbit than a retrograde orbit.
Fixing the orbit to be either exactly prograde or exactly retrograde,
we find $\Delta {\rm BIC} = {\rm
BIC}_{\rm retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the
retrograde case is ruled out on this basis.  More precise data will
however still be needed to make a precise obliquity measurement.
